---
title: "Conditioning diagnostics"
subtitle: "Operating Model - Anchovy 9a South"
author: "María José Zúñiga"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: false
    toc_depth: 3
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 5
)
```



```{r packages}
# 1. Load packages
library(FLBEIA)
library(FLCore)
library(dplyr)
library(tibble)
library(ggplot2)
library(stringr)
library(here)
library(readr)
library(zoo)
library(randtests)
library(changepoint)
library(strucchange)
library(broom)
library(patchwork)

set.seed(123)
```



```{r settings}
# 2. General settings
first_yr    <- 1989
last_obs_yr <- 2024

ss_rec <- 3   # Recruitment in Q3
ss_ssb <- 2   # SSB in Q2

Blim <- 4721
Bpa  <- 6561

stk_ane9aS_rds <- here("boot/data/stk_ane9aS.rds")
ss3_rds        <- here("boot/data/ss3_ane9aS.rds")

out_dir  <- here("outputs/recruitment_conditioning")
fig_dir  <- file.path(out_dir, "figures")
tab_dir  <- file.path(out_dir, "tables")
data_dir <- here("data/processed")

dir.create(fig_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(tab_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(data_dir, recursive = TRUE, showWarnings = FALSE)
```



```{r helper-functions}

# 3. Helper functions
find_value <- function(df, patterns, label_col = "Label", value_col = "Value") {
  if (!all(c(label_col, value_col) %in% names(df))) {
    stop("Columns ", label_col, " and/or ", value_col, " not found.")}

  pattern <- paste(patterns, collapse = "|")
  out <- df |> filter(str_detect(.data[[label_col]], regex(pattern, ignore_case = TRUE)))

  if (nrow(out) == 0) return(NA_real_)
  as.numeric(out[[value_col]][1])
}

ss3_bh <- function(SSB, R0, h, SSB0) {(4 * h * R0 * SSB) / (SSB0 * (1 - h) + SSB * (5 * h - 1))}
om_bh <- function(SSB, a, b) {a * SSB / (b + SSB)}

make_lognormal_bands <- function(rec_med, sigmaR) {
  tibble(
    rec_lower_95 = rec_med * exp(-0.5 * sigmaR^2 - 2.00 * sigmaR),
    rec_upper_95 = rec_med * exp(-0.5 * sigmaR^2 + 2.00 * sigmaR),
    rec_lower_80 = rec_med * exp(-0.5 * sigmaR^2 - 1.28 * sigmaR),
    rec_upper_80 = rec_med * exp(-0.5 * sigmaR^2 + 1.28 * sigmaR))}

save_plot <- function(plot, filename, width = 8, height = 5) {
  ggsave(filename = file.path(fig_dir, filename),
    plot = plot, width = width, height = height, dpi = 300)
}

safe_box_test <- function(x, lag) {
  out <- Box.test(x, lag = lag, type = "Ljung-Box")
  tibble(test = paste0("Ljung-Box lag ", lag),
         statistic = unname(out$statistic),
         p_value = out$p.value)
}

calc_skewness <- function(x) {
  x <- x[is.finite(x)]
  mean((x - mean(x))^3) / sd(x)^3}

calc_kurtosis <- function(x) {
  x <- x[is.finite(x)]
  mean((x - mean(x))^4) / sd(x)^4}

```



```{r load-stock}
# Load stock and define biological timing
ane_stock <- readRDS(stk_ane9aS_rds)

name(ane_stock) <- "ANE"
desc(ane_stock) <- "WGHANSA2025 assessment output"
ane_stock@harvest@units <- "f"
```


```{r biological-timing}
## Define mortality and maturity timing
fspwn <- harvest.spwn(ane_stock)
fspwn[] <- 0
fspwn[, , , "1", ] <- 1
harvest.spwn(ane_stock) <- fspwn

mspwn <- m.spwn(ane_stock)
mspwn[] <- 0
mspwn[, , , "1", ] <- 1
m.spwn(ane_stock) <- mspwn

mat(ane_stock)[ac(0), ] <- 0
mat(ane_stock)[, , , 1] <- 0
mat(ane_stock)[, , , 3] <- 0
mat(ane_stock)[, , , 4] <- 0

range(ane_stock, "minfbar") <- 3
range(ane_stock, "maxfbar") <- 3
```



```{r create-flbiol}
## Create FLBiol object
ane_biol <- FLBiol(
  n = stock.n(ane_stock),
  wt = stock.wt(ane_stock),
  m = m(ane_stock),
  spwn = m.spwn(ane_stock),
  mat = mat(ane_stock),
  fec = predictModel(
    FLQuants(fec = ane_stock@mat * 0 + 1),
    model = ~ fec),
  name = "ANE",
  desc = ane_stock@desc,
  range = ane_stock@range)

units(ane_biol)$m <- ""
units(fec(ane_biol)) <- ""
units(mat(ane_biol)) <- ""

ane_biol <- window(ane_biol, start = first_yr, end = last_obs_yr)
```



```{r extract-rec-ssb}
# Extract historical recruitment and SSB
rec_h <- rec(ane_biol)[, ac(first_yr:last_obs_yr), , ss_rec, drop = FALSE]
ssb_h <- ssb(ane_biol)[, ac(first_yr:last_obs_yr), , ss_ssb, drop = FALSE]

rec_fit <- iter(rec_h, 1)
ssb_fit <- iter(ssb_h, 1)

sr_data <- tibble(
  year = as.numeric(dimnames(rec_fit)$year),
  ssb  = as.numeric(ssb_fit),
  rec  = as.numeric(rec_fit)
) |>
  filter(is.finite(ssb), is.finite(rec), ssb > 0, rec > 0)

write_csv(sr_data, file.path(tab_dir, "00_historical_ssb_recruitment.csv"))

# head(sr_data)
```

# Fit stock-recruitment relationships

Three alternative stock–recruitment models were fitted to the historical SSB–recruitment observations: Beverton–Holt, Ricker and segmented regression (Figure X). The segmented model produced the lowest Akaike Information Criterion (AIC = -68.91), followed by the Beverton–Holt model (AIC = -68.84) and the Ricker model (AIC = -67.83) (Table X). The difference in AIC between the segmented and Beverton–Holt models was 0.08, whereas the Ricker model differed by 1.08 AIC units. 

The Beverton–Holt and segmented models produced similar recruitment estimates across most of the observed spawning biomass range. The Ricker model predicted lower recruitment at higher spawning biomass, reflecting an overcompensatory response (Figure X).


*Although the segmented model achieved the lowest AIC, its support relative to the Beverton–Holt model was negligible ($\Delta$AIC = 0.08). According to commonly used information-theoretic criteria, models differing by less than two AIC units receive similar empirical support. Consequently, model selection cannot be based on AIC alone.*

*The Beverton–Holt model was selected as the reference stock–recruitment relationship for the Operating Model because it provides a compensatory recruitment response that is consistent with the assessment model implemented in SS3, while requiring fewer parameters than the segmented model. The Ricker model was not retained as the reference model because it assumes overcompensation at high spawning biomass, a pattern that is not required to explain the historical observations.*


```{r fit-bh, include=FALSE}
## Beverton-Holt, Segmented and Ricker fits
mod_bh_om <- FLSR(rec = rec_fit, ssb = ssb_fit, model = bevholt)
fit_bh_om <- fmle(mod_bh_om)

mod_seg <- FLSR(rec = rec_fit, ssb = ssb_fit, model = "segreg")
fit_seg <- fmle(mod_seg)

mod_ricker <- FLSR(rec = rec_fit, ssb = ssb_fit, model = ricker)
fit_ricker <- fmle(mod_ricker)

```



```{r}
# Comparación de modelos
sr_model_comparison <- tibble(
  model = c("Beverton-Holt", "Ricker", "Segmented"),
  logLik = c(
    as.numeric(logLik(fit_bh_om)),
    as.numeric(logLik(fit_ricker)),
    as.numeric(logLik(fit_seg))),
  AIC = c(AIC(fit_bh_om), AIC(fit_ricker), AIC(fit_seg))) |>
  mutate(delta_AIC = AIC - min(AIC, na.rm = TRUE))

write_csv(sr_model_comparison,
  file.path(tab_dir, "31_SR_model_comparison_BH_Ricker_segmented.csv"))

knitr::kable(sr_model_comparison,digits = 3,
  caption = "Comparison of alternative stock-recruitment models fitted to historical SSB-recruitment observations.")
```


```{r}
hist_resid_om <- as.numeric(residuals(fit_bh_om))
hist_resid_om <- hist_resid_om[is.finite(hist_resid_om)]

sigmaR_om <- sd(hist_resid_om, na.rm = TRUE)

om_pars <- as.numeric(params(fit_bh_om))
names(om_pars) <- dimnames(params(fit_bh_om))$params

a_om <- om_pars["a"]
b_om <- om_pars["b"]

ssb_grid <- seq(min(sr_data$ssb, na.rm = TRUE) * 0.1, max(sr_data$ssb, na.rm = TRUE) * 1.2,length.out = 200)

om_curve <- tibble(ssb = ssb_grid, rec_med = om_bh(ssb_grid, a_om, b_om),model = "Beverton-Holt")
ricker_curve <- tibble(ssb = ssb_grid,rec_med = as.numeric(predict(fit_ricker, ssb = FLQuant(ssb_grid))), model = "Ricker")
seg_curve <- tibble(ssb = ssb_grid, rec_med = as.numeric(predict(fit_seg, ssb = FLQuant(ssb_grid))), model = "Segmented")

```



```{r}
p_sr_candidates <- ggplot() +
  geom_point(data = sr_data, aes(ssb, rec), size = 2, alpha = 0.7) +
  geom_line(data = om_curve, aes(ssb, rec_med, colour = "Beverton-Holt"), linewidth = 1.1) +
  geom_line(data = ricker_curve, aes(ssb, rec_med, colour = "Ricker"), linewidth = 1.1) +
  geom_line(data = seg_curve, aes(ssb, rec_med, colour = "Segmented"), linewidth = 1.1, linetype = "dashed") +
  geom_vline(xintercept = Blim, linetype = "dotted") +
  geom_vline(xintercept = Bpa, linetype = "dashed") +
  annotate("text", x = Blim, y = -Inf, label = "Blim", vjust = -0.6, size = 3, fontface = "italic") +
  annotate("text", x = Bpa, y = -Inf, label = "Bpa", vjust = -0.6, size = 3, fontface = "italic") +
  theme_bw() +
  labs( title = "Alternative stock-recruitment relationships",
    subtitle = "Beverton-Holt, Ricker and segmented models fitted to historical SSB-recruitment data",
    x = "SSB", y = "Recruitment", colour = "Model")

p_sr_candidates

save_plot(p_sr_candidates, "01_candidate_SR_models_BH_Ricker_segmented.png",width = 9, height = 6)
```

# Comparison between OM and SS3 Beverton–Holt relationships

The Beverton–Holt stock–recruitment relationship fitted in the Operating Model was compared with the Beverton–Holt relationship estimated in the SS3 assessment. The OM relationship was fitted directly to the historical SSB–recruitment pairs using the FLSR framework, whereas the SS3 relationship was derived from the assessment parameters \(R_0\), steepness and unfished spawning biomass.

Across the observed biomass range, the OM predicted lower recruitment than SS3 on average, with a mean difference of 4.1%. The mean ratio between OM and SS3 predictions was 0.96, with values ranging from 0.92 to 0.98. Recruitment residuals from the two relationships were highly correlated (r = 0.999), and the mean absolute difference between residuals was 0.042 log units.

<!-- The two Beverton–Holt relationships showed similar recruitment predictions across most of the observed biomass range. The SS3 curve predicted higher recruitment at low spawning biomass, while both curves approached a similar recruitment level at higher SSB. The residual standard deviation from the OM fit was 0.28, compared with $\sigma_R$=0.33 from SS3. -->


```{r load-ss3}
#Residuos y sigmaR Beverton & Holt
hist_resid_om <- as.numeric(residuals(fit_bh_om))
hist_resid_om <- hist_resid_om[is.finite(hist_resid_om)]

sigmaR_om <- sd(hist_resid_om, na.rm = TRUE)

## parámetros de Beverton & Holt OM
om_pars <- as.numeric(params(fit_bh_om))
names(om_pars) <- dimnames(params(fit_bh_om))$params

a_om <- om_pars["a"]
b_om <- om_pars["b"]

# load SS3 parameter
ss3rep <- readRDS(ss3_rds)
pars <- ss3rep$parameters
dq   <- ss3rep$derived_quants

lnR0_ss3 <- find_value(pars,c("SR_LN\\(R0\\)", "SR_Ln\\(R0\\)", "SR_log\\(R0\\)", "SR_LN_R0", "SR_lnR0"))

R0_ss3 <- exp(lnR0_ss3)
h_ss3 <- find_value(pars,c("SR_BH_steep", "SR_steep", "steep"))

sigmaR_ss3 <- find_value(pars,c("SR_sigmaR", "SR_sigma", "sigmaR", "SigmaR", "rec_sigma", "ln\\(R\\)_sigma"))

if (!is.finite(sigmaR_ss3)) {
  sigmaR_ss3 <- sigmaR_om
  sigmaR_source <- "OM residual SD"
} else {sigmaR_source <- "SS3"}

SSB0_ss3 <- find_value(dq,c("SSB_unfished", "SSB_Virgin", "SSB_virgin", "SSB_Bzero", "SSB_0", "SSB_unf"))

ss3_curve <- tibble(ssb = ssb_grid,rec_med = ss3_bh(ssb_grid, R0_ss3, h_ss3, SSB0_ss3),model = "Beverton-Holt SS3")

sr_parameter_table <- tibble(
  parameter = c("a_OM", "b_OM", "R0_SS3", "h_SS3","SSB0_SS3"),
  value = c(a_om, b_om, R0_ss3, h_ss3, SSB0_ss3),
  source = c("FLSR OM fit", "FLSR OM fit", "SS3", "SS3","SS3"))

write_csv(sr_parameter_table, file.path(tab_dir, "01_SR_parameter_table.csv"))

knitr::kable(sr_parameter_table,digits = 3,caption = "Stock-recruitment parameters used to compare the OM and SS3 Beverton-Holt relationships.")
```




```{r compare-om-ss3}
sr_obs <- sr_data |>
  mutate(
    rec_pred_om  = om_bh(ssb, a_om, b_om),
    rec_pred_ss3 = ss3_bh(ssb, R0_ss3, h_ss3, SSB0_ss3),
    resid_om     = log(rec) - log(rec_pred_om),
    resid_ss3    = log(rec) - log(rec_pred_ss3),
    ratio_om_ss3 = rec_pred_om / rec_pred_ss3,
    diff_percent = 100 * (rec_pred_om - rec_pred_ss3) / rec_pred_ss3)

compare_summary <- sr_obs |>
  summarise(
    n_years = dplyr::n(),
    mean_ratio_om_ss3 = mean(ratio_om_ss3, na.rm = TRUE),
    min_ratio_om_ss3 = min(ratio_om_ss3, na.rm = TRUE),
    max_ratio_om_ss3 = max(ratio_om_ss3, na.rm = TRUE),
    mean_diff_percent = mean(diff_percent, na.rm = TRUE),
    mean_abs_percent_diff = mean(abs(diff_percent), na.rm = TRUE),
    cor_residuals_om_ss3 = cor(resid_om, resid_ss3, use = "complete.obs"),
    mean_abs_residual_difference = mean(abs(resid_om - resid_ss3), na.rm = TRUE))

write_csv(sr_obs, file.path(tab_dir, "02_SR_observed_OM_vs_SS3.csv"))
write_csv(compare_summary, file.path(tab_dir, "03_OM_vs_SS3_summary.csv"))

# knitr::kable(
#   t(compare_summary),
#   digits = 3,
#   caption = "Summary comparison between OM and SS3 Beverton-Holt predictions."
# )
```


```{r plot-sr, fig.cap="Stock-recruitment relationship. Observed recruitment, OM Beverton-Holt fit, SS3 relationship  comparison"}
# 10. Plot SR relationship
p_sr_compare <- ggplot() +
  geom_point(data = sr_data,aes(ssb, rec),size = 2,alpha = 0.8) +
  geom_line(data = om_curve,aes(ssb, rec_med, colour = "OM Beverton-Holt"),linewidth = 1) +
  geom_line(data = ss3_curve,aes(ssb, rec_med, colour = "SS3 Beverton-Holt"),linewidth = 1,linetype = "dashed") +
  # geom_line(data = ricker_curve,aes(ssb, rec_med, colour = "Ricker"),linewidth = 1,linetype = "dotdash") +
  geom_vline(xintercept = Blim, linetype = "dotted") +
  geom_vline(xintercept = Bpa, linetype = "dashed") +
  annotate("text", x = Blim, y = -Inf, label = "Blim", vjust = -0.6, size = 3, fontface = "italic") +
  annotate("text", x = Bpa, y = -Inf, label = "Bpa", vjust = -0.6, size = 3, fontface = "italic") +
  theme_bw() +
  labs(title = "", x = "SSB",y = "Recruitment",colour = "Model")

p_sr_compare

save_plot(p_sr_compare, "02_SR_curve_OM_vs_SS3_sigmaR.png")
```


## Recruitment residual diagnostics

Recruitment residuals from the OM Beverton–Holt fit were examined to describe their distribution, temporal pattern, relationship with SSB and consistency with the recruitment process variability used in the OM.

The residuals were centred around zero, with a mean close to zero and an empirical standard deviation of 0.28. The value of \(\sigma_R\) extracted from the SS3 assessment was 0.33.

Thirty-five of the 36 historical residuals (97.2%) were within the approximate 95% prediction interval defined by \(\sigma_R = 0.33\). One residual was below the lower prediction limit and no residuals exceeded the upper limit.

The residual distribution showed slight negative skewness (-0.41) and kurtosis close to 3 (3.07). The Shapiro–Wilk test did not reject normality (W = 0.95, p = 0.08). Ljung–Box tests at lags 1, 3 and 5 were not significant (p = 0.46, 0.33 and 0.18, respectively).

Tail diagnostics showed that the observed lower-tail quantiles were less negative than expected under a normal distribution with \(\sigma_R = 0.33\). The observed 1%, 5% and 10% quantiles were -0.69, -0.37 and -0.28, compared with expected values of -0.82, -0.60 and -0.48. The observed 95% and 99% upper-tail quantiles were 0.39 and 0.41, compared with expected values of 0.49 and 0.71.

```{r residual-diagnostics}
rec_diag <- sr_obs |>
  transmute(
    year,
    ssb,
    rec_obs = rec,
    rec_pred_om,
    rec_pred_ss3,
    rec_resid_om = resid_om,
    rec_resid_ss3 = resid_ss3,
    rec_mult_om = exp(resid_om),
    sigmaR_lower_95 = -0.5 * sigmaR_ss3^2 - 2.00 * sigmaR_ss3,
    sigmaR_upper_95 = -0.5 * sigmaR_ss3^2 + 2.00 * sigmaR_ss3,
    below_sigmaR_95 = rec_resid_om < sigmaR_lower_95,
    above_sigmaR_95 = rec_resid_om > sigmaR_upper_95
  ) |>
  mutate(
    inside_sigmaR_95 = !below_sigmaR_95 & !above_sigmaR_95
  )

write_csv(rec_diag,file.path(tab_dir, "04_recruitment_residual_diagnostics_by_year.csv"))
```


```{r residual-summary}
resid_summary <- tibble(
  n_years = nrow(rec_diag),
  mean_resid = mean(rec_diag$rec_resid_om, na.rm = TRUE),
  sd_resid = sd(rec_diag$rec_resid_om, na.rm = TRUE),
  sigmaR_used = sigmaR_ss3,
  sigmaR_source = sigmaR_source,
  skewness = calc_skewness(rec_diag$rec_resid_om),
  kurtosis = calc_kurtosis(rec_diag$rec_resid_om),
  q01 = quantile(rec_diag$rec_resid_om, 0.01, na.rm = TRUE),
  q05 = quantile(rec_diag$rec_resid_om, 0.05, na.rm = TRUE),
  q10 = quantile(rec_diag$rec_resid_om, 0.10, na.rm = TRUE),
  median = median(rec_diag$rec_resid_om, na.rm = TRUE),
  q90 = quantile(rec_diag$rec_resid_om, 0.90, na.rm = TRUE),
  q95 = quantile(rec_diag$rec_resid_om, 0.95, na.rm = TRUE),
  q99 = quantile(rec_diag$rec_resid_om, 0.99, na.rm = TRUE),
  min_resid = min(rec_diag$rec_resid_om, na.rm = TRUE),
  max_resid = max(rec_diag$rec_resid_om, na.rm = TRUE),
  prop_inside_sigmaR_95 = mean(rec_diag$inside_sigmaR_95, na.rm = TRUE),
  prop_below_sigmaR_95 = mean(rec_diag$below_sigmaR_95, na.rm = TRUE),
  prop_above_sigmaR_95 = mean(rec_diag$above_sigmaR_95, na.rm = TRUE)
)

resid_tests <- bind_rows(
  safe_box_test(rec_diag$rec_resid_om, lag = 1),
  safe_box_test(rec_diag$rec_resid_om, lag = 3),
  safe_box_test(rec_diag$rec_resid_om, lag = 5),
  {
    sw <- shapiro.test(rec_diag$rec_resid_om)
    tibble(
      test = "Shapiro-Wilk normality",
      statistic = unname(sw$statistic),
      p_value = sw$p.value
    )
  }
)

tail_diagnostics <- tibble(
  probability = c(0.01, 0.05, 0.10, 0.50, 0.90, 0.95, 0.99),
  observed_residual = as.numeric(
    quantile(rec_diag$rec_resid_om, probability, na.rm = TRUE)
  ),
  normal_expected = qnorm(
    probability,
    mean = -0.5 * sigmaR_ss3^2,
    sd = sigmaR_ss3
  )
) |>
  mutate(
    difference_observed_minus_expected =
      observed_residual - normal_expected
  )

```


```{r plot-residual-time-series}
p_resid_ts <- ggplot(rec_diag, aes(year, rec_resid_om)) +
  geom_rect(aes(xmin = -Inf,xmax = Inf,ymin = sigmaR_lower_95,ymax = sigmaR_upper_95),
    inherit.aes = FALSE, fill = "grey80",alpha = 0.15) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_line(linewidth = 0.5) +
  geom_point(size = 2.2) +
  theme_bw() +
  labs(x = "Year",y = "Recruitment residual",colour = NULL)

save_plot(p_resid_ts, "03_OM_residuals_time_series_sigmaR.png")
```


```{r plot-residual-distribution}
p_resid_hist <- ggplot(rec_diag, aes(rec_resid_om)) +
  geom_histogram(aes(y = after_stat(density)),bins = 12,fill = "grey80",colour = "black") +
  stat_function(fun = dnorm, args = list(mean = -0.5 * sigmaR_ss3^2, sd = sigmaR_ss3),
    linewidth = 1,linetype = "dashed") +
  theme_bw() +
  labs(x = "Recruitment residual",y = "Density")

save_plot(p_resid_hist, "04_OM_residual_distribution_vs_sigmaR.png")
```

```{r plot-qq}
p_qq <- ggplot(rec_diag, aes(sample = rec_resid_om)) +
  stat_qq() + stat_qq_line() +
  theme_bw() +
  labs(title = "(b) Normal Q-Q plot",x = "Theoretical quantiles",y = "Observed residual quantiles")
# p_qq
save_plot(p_qq, "05_OM_residual_QQplot.png")
```

```{r plot-residuals-ssb}
p_resid_ssb <- ggplot(rec_diag, aes(ssb, rec_resid_om)) +
  geom_rect(aes(xmin = -Inf,xmax = Inf,ymin = sigmaR_lower_95,ymax = sigmaR_upper_95),
    inherit.aes = FALSE,fill = "grey80",alpha = 0.35) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = Blim, linetype = "dotted") +
  geom_vline(xintercept = Bpa, linetype = "dashed") +
  geom_point(size = 2) +
  geom_smooth(method = "loess", se = TRUE, linewidth = 0.8) +
  theme_bw() +
  labs(title = "Residuals vs SSB",subtitle = "LOESS smoother",
       x = "SSB",y = "Recruitment residual",colour = NULL)

save_plot(p_resid_ssb, "06_OM_residuals_vs_SSB_sigmaR.png")
```



```{r tail-diagnostics}
tail_diagnostics <- tibble(
  probability = c(0.01, 0.05, 0.10, 0.50, 0.90, 0.95, 0.99),
  observed_residual = as.numeric(
    quantile(rec_diag$rec_resid_om, probability, na.rm = TRUE)
  ),
  normal_expected = qnorm(
    probability,
    mean = -0.5 * sigmaR_ss3^2,
    sd = sigmaR_ss3
  )) |>
  mutate(
    difference_observed_minus_expected =
      observed_residual - normal_expected
  )
write_csv(tail_diagnostics, file.path(tab_dir, "07_tail_diagnostics.csv"))

# tail_diagnostics

tail_severity <- tail_diagnostics |> filter(probability %in% c(0.01, 0.05, 0.10)) |>
  mutate(severity_ratio = abs(observed_residual) / abs(normal_expected),
         interpretation = case_when(severity_ratio > 1.25 ~ "historical tail more extreme",
                                    severity_ratio < 0.75 ~ "historical tail less extreme",
                                    TRUE ~ "similar to lognormal assumption"))

write_csv(tail_severity, file.path(tab_dir, "18_tail_severity.csv"))


# t(tail_severity)
```

```{r plot-tail-diagnostics}
p_tail <- ggplot(tail_diagnostics, aes(normal_expected, observed_residual)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  geom_point(size = 2) +
  geom_text(aes(label = probability), nudge_y = 0.05, size = 3) +
  theme_bw() +
  labs(
    title = "Tail diagnostics",
    subtitle = "Observed recruitment residual quantiles compared with expected normal quantiles",
    x = "Expected residual quantile",
    y = "Observed residual quantile"
  )

save_plot(p_tail, "07_tail_diagnostics.png")
```


```{r fig.height=6,fig.width=6}
panel_theme <- theme_bw() +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 10, face = "bold"),
    plot.subtitle = element_text(size = 8),
    axis.title = element_text(size = 9),
    axis.text = element_text(size = 8),
    plot.margin = margin(5, 5, 5, 5))

p_resid_ts2 <- p_resid_ts +
  labs(title = "(a) Residuals through time",subtitle = "Grey band: approximate 95% prediction interval") +
  panel_theme

p_resid_hist2 <- p_resid_hist +
  labs(title = "(b) Residual distribution",subtitle = "Dashed line: expected normal distribution") +
  panel_theme

p_resid_ssb2 <- p_resid_ssb +
  labs(title = "(c) Residuals vs SSB",subtitle = "LOESS smoother") +
  panel_theme

p_qq2 <- p_qq +
  labs(title = "(d) Normal Q-Q plot",subtitle = NULL) +
  panel_theme

residuals_diagnostic_plot <-
  (p_resid_ts2 | p_resid_hist2) /
  (p_resid_ssb2 | p_qq2) +
  plot_annotation(
    title = "Recruitment residual diagnostics",
    subtitle = "Temporal pattern, distribution, biomass-related structure and normality",
    theme = theme(
      plot.title = element_text(face = "bold", size = 15),
      plot.subtitle = element_text(size = 10)))

residuals_diagnostic_plot
```

## Autocorrelation diagnostics

The ACF and PACF were used to examine temporal dependence in the OM recruitment residuals. All ACF coefficients up to lag 15 were within the approximate 95% limits. In the PACF, one coefficient at lag 5 slightly exceeded the 95% limit, while the remaining lags were within the interval.

An AR(1) model fitted to the residuals produced a low autoregressive coefficient (`ar1 = 0.12`) that was not statistically different from zero (*p* = 0.48). These results did not indicate a consistent autoregressive pattern in the recruitment residuals.

```{r autocorrelation}
resid_vec <- rec_diag$rec_resid_om
resid_vec <- resid_vec[is.finite(resid_vec)]

acf_obj  <- acf(resid_vec, plot = FALSE)
pacf_obj <- pacf(resid_vec, plot = FALSE)

acf_table <- tibble(
  lag = as.numeric(acf_obj$lag[-1]),
  acf = as.numeric(acf_obj$acf[-1]),
  ci95 = 1.96 / sqrt(length(resid_vec)),
  significant = abs(acf) > ci95)

pacf_table <- tibble(
  lag = as.numeric(pacf_obj$lag),
  pacf = as.numeric(pacf_obj$acf),
  ci95 = 1.96 / sqrt(length(resid_vec)),
  significant = abs(pacf) > ci95)

ar1_fit <- arima(resid_vec, order = c(1, 0, 0), include.mean = TRUE)

ar1_table <- tibble(
  parameter = names(ar1_fit$coef),
  estimate = as.numeric(ar1_fit$coef),
  se = sqrt(diag(ar1_fit$var.coef)),
  z = estimate / se,
  p_value = 2 * pnorm(abs(z), lower.tail = FALSE))

write_csv(acf_table, file.path(tab_dir, "10_ACF_recruitment_residuals.csv"))
write_csv(pacf_table, file.path(tab_dir, "11_PACF_recruitment_residuals.csv"))
write_csv(ar1_table, file.path(tab_dir, "12_AR1_recruitment_residuals.csv"))

```

```{r}
# knitr::kable(round(acf_table, 3),caption = "Autocorrelation coefficients of recruitment residuals.")
# knitr::kable(round(pacf_table, 3),caption = "Partial autocorrelation coefficients of recruitment residuals.")
knitr::kable(ar1_table, caption = "Estimated parameters of the AR(1) model fitted to recruitment residuals.")
```

```{r}
op <- par(mfrow = c(1, 2),mar = c(4, 4, 3, 1))
acf(resid_vec,main = "ACF")
pacf(resid_vec,main = "PACF")
par(op)
```

## Change-point diagnostics

Change-point analyses were applied to the historical log-recruitment series to identify potential shifts in the mean or in both the mean and variance. Using the `PELT` algorithm with the `MBIC` penalty, no change-points were detected in either the mean-only model or the mean-and-variance model.

A complementary breakpoint analysis using `strucchange` was also applied to the log-recruitment series. No finite breakpoints were retained in the resulting breakpoint table.

Overall, these diagnostics did not identify discrete changes in the historical log-recruitment series over the 1989–2024 period.

Because no changepoints were detected in the historical log-recruitment series, the full period 1989–2024 was treated as a single recruitment period. Summary statistics were therefore calculated for the complete historical series.

No changepoints were detected in the log-recruitment series using the mean-only or mean-and-variance models. Consequently, the full period 1989–2024 was retained as a single recruitment period. Mean recruitment over this period was 8.10 million recruits, with a standard deviation of 2.19 million and a geometric mean of 7.80 million.

```{r changepoint}
log_rec   <- log(sr_data$rec)
years_rec <- sr_data$year

cpt_mean <- changepoint::cpt.mean(log_rec,method = "PELT",penalty = "MBIC")

cpt_meanvar <- changepoint::cpt.meanvar(log_rec,method = "PELT",penalty = "MBIC")

cpt_table <- tibble(method = c("mean", "mean_and_variance"),
  n_change_points = c(length(changepoint::cpts(cpt_mean)),length(changepoint::cpts(cpt_meanvar))),
  change_point_index = c(paste(changepoint::cpts(cpt_mean), collapse = "; "),paste(changepoint::cpts(cpt_meanvar), collapse = "; ")),
  change_point_year = c(paste(years_rec[changepoint::cpts(cpt_mean)], collapse = "; "),paste(years_rec[changepoint::cpts(cpt_meanvar)], collapse = "; ")))

bp <- strucchange::breakpoints(log_rec ~ 1)

bp_table <- tibble(
  method = "strucchange_breakpoints",
  break_index = bp$breakpoints,
  break_year = years_rec[bp$breakpoints]) |>
  filter(is.finite(break_index))

write_csv(cpt_table, file.path(tab_dir, "13_changepoint_recruitment.csv"))
write_csv(bp_table, file.path(tab_dir, "14_strucchange_breakpoints_recruitment.csv"))

knitr::kable(
  cpt_table,
  caption = "Change-point diagnostics applied to historical log recruitment."
)
```





```{r recruitment-regimes}
cp_idx <- changepoint::cpts(cpt_meanvar)
cp_years <- sr_data$year[cp_idx]

if (length(cp_idx) == 0) {
  
  regime_breaks <- c(
    min(sr_data$year),
    max(sr_data$year) + 1
  )
  
} else {
  
  regime_breaks <- c(
    min(sr_data$year),
    cp_years + 1,
    max(sr_data$year) + 1
  )
}

regime_info <- tibble(
  start_year = regime_breaks[-length(regime_breaks)],
  end_year   = regime_breaks[-1] - 1,
  regime_id  = seq_along(start_year)
) |>
  rowwise() |>
  mutate(
    mean_rec = mean(
      sr_data$rec[sr_data$year >= start_year & sr_data$year <= end_year],
      na.rm = TRUE
    ),
    sd_rec = sd(
      sr_data$rec[sr_data$year >= start_year & sr_data$year <= end_year],
      na.rm = TRUE
    ),
    geomean_rec = exp(mean(
      log(sr_data$rec[sr_data$year >= start_year & sr_data$year <= end_year]),
      na.rm = TRUE
    )),
    ymin = pmax(mean_rec - sd_rec, 0),
    ymax = mean_rec + sd_rec,
    regime_label = paste0(regime_id, ") ", start_year, "-", end_year)
  ) |>
  ungroup()

regime_diag_cp <- sr_data |>
  mutate(
    regime_id = findInterval(
      year,
      vec = regime_breaks,
      rightmost.closed = TRUE
    )) |>
  left_join(
    regime_info |> select(regime_id, regime_label, geomean_rec),
    by = "regime_id"
  ) |>
  mutate(
    regime_label = factor(regime_label, levels = regime_info$regime_label))

regime_cols <- setNames(
  c("#66C2A5", "#8DA0CB", "#FFD92F", "#E78AC3", "#FC8D62")[seq_len(nrow(regime_info))],
  regime_info$regime_label)

write_csv(regime_info, file.path(tab_dir, "40_recruitment_period_summary.csv"))
write_csv(regime_diag_cp, file.path(tab_dir, "41_recruitment_period_by_year.csv"))

# knitr::kable(
#   t(regime_info),
#   digits = 2,
#   caption = "Summary statistics for recruitment periods defined from detected changepoints.")
```

# Recruitment temporal diagnostics

Recruitment and SSB were examined through time to evaluate temporal structure in the historical series. Since no changepoints were detected in log recruitment, the full 1989–2024 period was treated as a single recruitment period. The shaded band in panel (a) represents the mean ± SD of recruitment over the full period, and the horizontal line indicates the geometric mean.

SSB was rescaled in panel (a) only for visual comparison with recruitment. The secondary axis converts the scaled values back to the original SSB scale.

Spearman correlations were calculated between recruitment and SSB at lags from -3 to +3 years. Negative lags indicate correlations between past recruitment and current SSB, whereas positive lags indicate correlations between current SSB and future recruitment.



```{r plot-recruitment-regimes, fig.width=10, fig.height=6}
ssb_scale <- max(regime_diag_cp$rec, na.rm = TRUE) /
  max(regime_diag_cp$ssb, na.rm = TRUE)

p_ts <- ggplot() +
  geom_rect(
    data = regime_info,
    aes(
      xmin = start_year - 0.5,
      xmax = end_year + 0.5,
      ymin = ymin,
      ymax = ymax,
      fill = regime_label
    ),
    alpha = 0.25,
    colour = NA
  ) +
  geom_line(
    data = regime_diag_cp,
    aes(year, rec),
    linewidth = 0.55
  ) +
  geom_point(
    data = regime_diag_cp,
    aes(year, rec, fill = regime_label),
    shape = 21,
    size = 2,
    colour = "black",
    alpha = 0.85
  ) +
  geom_segment(
    data = regime_info,
    aes(
      x = start_year,
      xend = end_year,
      y = geomean_rec,
      yend = geomean_rec
    ),
    linewidth = 1.15
  ) +
  geom_line(
    data = regime_diag_cp,
    aes(year, ssb * ssb_scale),
    linewidth = 0.55,
    linetype = "dashed"
  ) +
  geom_vline(
    xintercept = cp_years + 0.5,
    linetype = "dashed"
  ) +
  scale_fill_manual(values = regime_cols, drop = FALSE) +
  scale_y_continuous(
    name = "Recruitment",
    sec.axis = sec_axis(~ . / ssb_scale, name = "SSB")
  ) +
  theme_bw() +
  theme(
    legend.position = "top"
  ) +
  labs(
  title = "(a) Recruitment and spawning biomass through time",
  subtitle = "Single period defined from changepoint analysis; shaded area: mean ± SD recruitment",
  x = NULL,
  fill = "Period"
)

# p_ts

save_plot(p_ts,"27a_recruitment_and_SSB_regime_periods.png",width = 10,height = 6)
```





```{r cross-correlation-function}
calc_ccf_paper <- function(dat, max_lag = 3) {

  bind_rows(lapply(-max_lag:max_lag, function(lag) {

    if (lag < 0) {
      tmp <- tibble(ssb = dat$ssb[(1 - lag):nrow(dat)],rec = dat$rec[1:(nrow(dat) + lag)])
    } else if (lag > 0) {
      tmp <- tibble(ssb = dat$ssb[1:(nrow(dat) - lag)],rec = dat$rec[(1 + lag):nrow(dat)])
    } else {
      tmp <- tibble(ssb = dat$ssb,rec = dat$rec)
    }

    test <- cor.test(tmp$ssb,tmp$rec,method = "spearman",exact = FALSE)
    tibble(lag = lag,rho = unname(test$estimate),p_value = test$p.value,n = nrow(tmp))
  }))
}

ccf_paper <- calc_ccf_paper(sr_data, max_lag = 3)

write_csv(ccf_paper,file.path(tab_dir, "42_cross_correlation_recruitment_SSB.csv"))

# ccf_paper
```


```{r plot-cross-correlation, fig.width=9, fig.height=5}
ccf_plot <- ccf_paper |>
  mutate(
    lag_label = case_when(
      lag < 0 ~ paste0("R(t", lag, ") \u2192 SSB(t)"),
      lag == 0 ~ "SSB(t) \u2192 R(t)",
      lag > 0 ~ paste0("SSB(t) \u2192 R(t+", lag, ")")),
    lag_label = factor(lag_label, levels = lag_label),
    significant = p_value < 0.05)

ci95 <- 1.96 / sqrt(nrow(sr_data))

p_ccf <- ggplot(ccf_plot, aes(lag_label, rho)) +
  geom_col(fill = "grey45", width = 0.75) +
  geom_hline(yintercept = 0) +
  geom_hline(
    yintercept = c(-ci95, ci95),
    linetype = "dashed",
    colour = "blue") +
  geom_text(
    aes(label = ifelse(significant, "*", "")),
    vjust = -0.35,
    size = 5) +
  coord_cartesian(ylim = c(-1, 1)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 35, hjust = 1)) +
  labs(
    title = "(c) Cross-correlation between recruitment and SSB",
    subtitle = paste0(
      "Full period ", first_yr, "\u2013", last_obs_yr,
      ". Negative lags: past recruitment contributes to current SSB; ",
      "lag 0: SSB Q", ss_ssb, " to recruitment Q", ss_rec),
    x = NULL,
    y = expression(Spearman~rho))

save_plot(p_ccf,"27c_cross_correlation_recruitment_SSB.png",width = 9,height = 5)
```


```{r final-recruitment-diagnostic-figure, fig.width=15, fig.height=11}
p_sr <- ggplot(regime_diag_cp, aes(ssb, rec)) +
  geom_point(aes(fill = regime_label),shape = 21,size = 3,alpha = 0.85,colour = "black") +
  geom_line(data = om_curve,aes(x = ssb, y = rec_med, linetype = "Beverton-Holt"),inherit.aes = FALSE,linewidth = 1,colour = "black") +
  geom_line(data = ricker_curve,aes(x = ssb, y = rec_med, linetype = "Ricker"),inherit.aes = FALSE,linewidth = 1,colour = "black") +
  geom_vline(xintercept = Blim, linetype = "dotted") +
  geom_vline(xintercept = Bpa, linetype = "dashed") +
  annotate("text",x = Blim,y = -Inf,label = "Blim",vjust = -0.6,size = 3,fontface = "italic") +
  annotate("text",x = Bpa,y = -Inf,label = "Bpa",vjust = -0.6,size = 3,fontface = "italic") +
  scale_fill_manual(values = regime_cols, drop = FALSE) +
  scale_linetype_manual(values = c("Beverton-Holt" = "solid", "Ricker" = "dashed"),name = "SR model") +
  theme_bw() +
  theme(legend.position = "right") +
  labs(
  title = "(b) Stock-recruitment relationship",
  subtitle = "Observations correspond to the full 1989–2024 period",
  x = "SSB",
  y = "Recruitment",
  fill = "Period"
)

final_recruitment_diagnostic_plot <- p_ts / p_sr / p_ccf +
  plot_layout(heights = c(1.1, 1, 0.8)) +
  plot_annotation(
    title = "European anchovy 9a South recruitment diagnostics",
    caption = paste(
      "Common name: European anchovy;",
      "Species: Engraulis encrasicolus;",
      "Stock: 9a South, Gulf of Cádiz;",
      "Data source: SS3 assessment;",
      "Years:", paste0(first_yr, "\u2013", last_obs_yr), "."))

final_recruitment_diagnostic_plot

ggsave(filename = file.path(fig_dir, "27_recruitment_diagnostics_overview.png"),
  plot = final_recruitment_diagnostic_plot,
  width = 15,
  height = 11,
  dpi = 300)

ggsave(filename = file.path(fig_dir, "27_recruitment_diagnostics_overview.pdf"),
  plot = final_recruitment_diagnostic_plot,
  width = 15,
  height = 11)
```

# Recruitment productivity

Recruitment productivity, expressed as recruitment per unit of spawning stock biomass (R/SSB), showed substantial interannual variability throughout the historical period (Figure X). The 5-year moving average fluctuated between approximately 500 and 750 recruits per unit of SSB, with no persistent long-term increasing or decreasing trend. Recruitment exhibited a similar temporal pattern, with marked year-to-year fluctuations and a moving average ranging between approximately 6 and 9 million recruits.

Summary statistics were calculated for the full historical period (1989–2024) and for two recent subsets (2005–present and 2015–present) (Table X). Mean recruitment productivity ranged from 611 recruits per unit of SSB over the full time series to 655 recruits per unit of SSB during 2015–present. The corresponding geometric means ranged from 572 to 622 recruits per unit of SSB. The coefficient of variation of recruitment productivity was 0.37 over the full period, 0.40 for 2005–present and 0.34 for 2015–present.

Formal comparisons between the historical period and the recent period showed no statistically significant differences in recruitment productivity. Neither the Welch two-sample t-test (p = 0.48) nor the Wilcoxon rank-sum test (p = 0.45) detected significant differences between periods.

```{r productivity}
sr_data <- sr_data |>
  mutate(
    productivity = rec / ssb,
    log_productivity = log(productivity),
    productivity_roll5 = zoo::rollmean(productivity, k = 5, fill = NA, align = "center"),
    period_2015 = ifelse(year < 2015, "Historical", "Recent"))

p_prod <- ggplot(sr_data, aes(year, productivity)) +
  geom_line(alpha = 0.4) +
  geom_point(size = 1.8) +
  geom_line(aes(y = productivity_roll5), linewidth = 1.1) +
  theme_bw() +
  labs(
    title = "Recruitment productivity",
    subtitle = "Recruitment per unit of SSB, with 5-year moving average",
    x = "Year",
    y = "R / SSB")


save_plot(p_prod, "13_productivity_R_per_SSB_roll5.png")
```

```{r productivity-tests}
period_summary <- bind_rows(
  sr_data |>
    summarise(
      period = paste0(min(year), "-", max(year)),
      mean_R = mean(rec, na.rm = TRUE),
      geomean_R = exp(mean(log(rec), na.rm = TRUE)),
      cv_R = sd(rec, na.rm = TRUE) / mean(rec, na.rm = TRUE),
      mean_prod = mean(productivity, na.rm = TRUE),
      geomean_prod = exp(mean(log_productivity, na.rm = TRUE)),
      cv_prod = sd(productivity, na.rm = TRUE) / mean(productivity, na.rm = TRUE),
      n_years = dplyr::n()),
  sr_data |>
    filter(year >= 2005) |>
    summarise(
      period = "2005-present",
      mean_R = mean(rec, na.rm = TRUE),
      geomean_R = exp(mean(log(rec), na.rm = TRUE)),
      cv_R = sd(rec, na.rm = TRUE) / mean(rec, na.rm = TRUE),
      mean_prod = mean(productivity, na.rm = TRUE),
      geomean_prod = exp(mean(log_productivity, na.rm = TRUE)),
      cv_prod = sd(productivity, na.rm = TRUE) / mean(productivity, na.rm = TRUE),
      n_years = dplyr::n()),
  sr_data |>
    filter(year >= 2015) |>
    summarise(
      period = "2015-present",
      mean_R = mean(rec, na.rm = TRUE),
      geomean_R = exp(mean(log(rec), na.rm = TRUE)),
      cv_R = sd(rec, na.rm = TRUE) / mean(rec, na.rm = TRUE),
      mean_prod = mean(productivity, na.rm = TRUE),
      geomean_prod = exp(mean(log_productivity, na.rm = TRUE)),
      cv_prod = sd(productivity, na.rm = TRUE) / mean(productivity, na.rm = TRUE),
      n_years = dplyr::n()))

write_csv(period_summary, file.path(tab_dir, "20_productivity_by_period.csv"))



#----------------------------------------------------------------------------------
prod_ttest  <- t.test(productivity ~ period_2015, data = sr_data)
prod_wilcox <- wilcox.test(productivity ~ period_2015, data = sr_data)

productivity_tests <- tibble(
  test = c("Welch t-test", "Wilcoxon rank-sum test"),
  statistic = c(unname(prod_ttest$statistic), unname(prod_wilcox$statistic)),
  p_value = c(prod_ttest$p.value, prod_wilcox$p.value)
  # interpretation = ifelse(
  #   p_value < 0.05,
  #   "Evidence of difference between historical and recent productivity",
  #   "No evidence of difference between historical and recent productivity")
  )

write_csv(productivity_tests, file.path(tab_dir, "21_productivity_historical_vs_recent_tests.csv"))


```



```{r recruitment-rolling-mean}
sr_data <- sr_data |>
  mutate(rec_roll5 = zoo::rollmean(rec, k = 5, fill = NA, align = "center"))

p_roll <- ggplot(sr_data, aes(year)) +
  geom_line(aes(y = rec), alpha = 0.35) +
  geom_point(aes(y = rec), alpha = 0.5, size = 1.5) +
  geom_line(aes(y = rec_roll5), linewidth = 1.2) +
  theme_bw() +
  labs(
    title = "Historical recruitment with 5-year moving average",
    subtitle = "Visual diagnostic for persistent productivity changes",
    x = "Year",
    y = "Recruitment")

save_plot(p_roll, "10_recruitment_rolling_mean_5yr.png")
```

```{r}
five_year_moving_plot <- p_prod /p_roll +
  plot_annotation(
    title = "",
    theme = theme(
      plot.title = element_text(face = "bold", size = 15),
      plot.subtitle = element_text(size = 10)))
five_year_moving_plot
```

```{r}
knitr::kable(t(period_summary))
knitr::kable(productivity_tests)
```

# Density-dependence diagnostics

Recruitment productivity declined with increasing spawning stock biomass (Figure X). A linear regression of log recruitment productivity (log(R/SSB)) against log spawning stock biomass (log(SSB)) yielded a negative slope of -0.866 (SE = 0.171), which differed significantly from zero (t = -5.07, p < 0.001). The fitted model explained 43.0% of the observed variation in recruitment productivity (adjusted $R^2$ = 0.413). The fitted regression line and its 95% confidence interval are shown in Figure X.

*The observed decline in recruitment productivity with increasing spawning biomass is consistent with compensatory recruitment dynamics. This pattern supports the use of a compensatory stock-recruitment formulation, such as the Beverton–Holt model, in the Operating Model.*

```{r density-dependence}
sr_data <- sr_data |>
  mutate(
    RPS = rec / ssb,
    log_RPS = log(rec / ssb),
    log_SSB = log(ssb))

comp_lm <- lm(log_RPS ~ log_SSB, data = sr_data)

comp_test <- broom::tidy(comp_lm)
comp_test2<-broom::glance(comp_lm) |>
  dplyr::select(r.squared, adj.r.squared)

write_csv(comp_test, file.path(tab_dir, "27_compensation_log_RPS_vs_SSB.csv"))

# comp_test
```

```{r plot-density-dependence}
p_rps_ssb <- ggplot(sr_data, aes(ssb, RPS)) +
  geom_vline(xintercept = Blim, linetype = "dotted") +
  geom_vline(xintercept = Bpa, linetype = "dashed") +
  geom_point(size = 2) +
  geom_smooth(method = "loess", se = TRUE, linewidth = 0.8) +
  theme_bw() +
  labs(
    title = "Recruitment per spawner vs SSB",
    subtitle = "Diagnostic for compensatory recruitment dynamics",
    x = "SSB",
    y = "Recruitment / SSB")

p_log_rps_ssb <- ggplot(sr_data, aes(log_SSB, log_RPS)) +
  # geom_vline(xintercept = Blim, linetype = "dotted") +
  # geom_vline(xintercept = Bpa, linetype = "dashed") +
  geom_point(size = 2) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.8) +
  theme_bw() +
  labs(
    title = "log(R/SSB) vs log(SSB)",
    subtitle = "Linear regression with 95% confidence interval",
    x = "log(SSB)",
    y = "log(Recruitment / SSB)"
  )

# p_rps_ssb
p_log_rps_ssb

save_plot(p_rps_ssb, "15_RPS_vs_SSB_compensation.png")
save_plot(p_log_rps_ssb, "16_log_RPS_vs_SSB_compensation.png")
```

# Recruitment residual magnitude and spawning biomass

The magnitude of recruitment residuals showed no apparent relationship with spawning stock biomass (Figure X). A Spearman rank correlation between the absolute recruitment residuals and SSB was weak ($\rho$ = -0.09) and not statistically significant (p = 0.61). The fitted LOESS smoother remained approximately constant across the observed biomass range, with no evident increase or decrease in residual magnitude.

The distribution of absolute recruitment residuals was also similar between years classified as low and high spawning biomass (Figure X). Mean absolute residuals were 0.232 and 0.218 for the low- and high-biomass classes, respectively, while median absolute residuals were 0.169 and 0.191. A Wilcoxon rank-sum test detected no significant difference between biomass classes (W = 166, p = 0.91).

*The available historical data provide no evidence that the magnitude of recruitment deviations varies with spawning stock biomass. Therefore, recruitment process variability was represented in the Operating Model using a constant value of $\sigma_R$, rather than allowing variance to depend on stock size.*

```{r abs-residuals}
rec_diag <- rec_diag |>
  mutate(
    abs_rec_resid_om = abs(rec_resid_om),
    biomass_class = ifelse(
      ssb < median(ssb, na.rm = TRUE),
      "Low SSB",
      "High SSB"))

abs_resid_spearman <- cor.test(
  rec_diag$abs_rec_resid_om,
  rec_diag$ssb,
  method = "spearman",
  exact = FALSE)

abs_resid_spearman_table <- tibble(
  test = "Spearman correlation",
  variable_x = "|recruitment residual|",
  variable_y = "SSB",
  statistic = unname(abs_resid_spearman$statistic),
  estimate = unname(abs_resid_spearman$estimate),
  p_value = abs_resid_spearman$p.value)

abs_resid_by_biomass <- rec_diag |>
  group_by(biomass_class) |>
  summarise(
    n = dplyr::n(),
    mean_abs_resid = mean(abs_rec_resid_om, na.rm = TRUE),
    median_abs_resid = median(abs_rec_resid_om, na.rm = TRUE),
    sd_abs_resid = sd(abs_rec_resid_om, na.rm = TRUE),
    .groups = "drop"
  )
abs_resid_wilcox <- wilcox.test(abs_rec_resid_om ~ biomass_class, data = rec_diag)

abs_resid_wilcox_table <- tibble(
  test = "Wilcoxon rank-sum test",
  comparison = "Absolute recruitment residuals by biomass class",
  statistic = unname(abs_resid_wilcox$statistic),
  p_value = abs_resid_wilcox$p.value)

write_csv(abs_resid_spearman_table, file.path(tab_dir, "24_abs_residuals_vs_SSB_spearman.csv"))
write_csv(abs_resid_by_biomass, file.path(tab_dir, "25_abs_residuals_by_biomass_class.csv"))
write_csv(abs_resid_wilcox_table, file.path(tab_dir, "26_abs_residuals_by_biomass_class_wilcox.csv"))

abs_resid_spearman_table
abs_resid_by_biomass
abs_resid_wilcox_table
```

```{r plot-abs-residuals}
p_abs_resid_ssb <- ggplot(rec_diag, aes(ssb, abs_rec_resid_om)) +
  geom_vline(xintercept = Blim, linetype = "dotted") +
  geom_vline(xintercept = Bpa, linetype = "dashed") +
  geom_point(size = 2) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.8) +
  theme_bw() +
  labs(
    title = "Absolute recruitment residuals vs SSB",
    subtitle = "Diagnostic for stock-size-dependent recruitment variability",
    x = "SSB",
    y = "|Recruitment residual|")

p_abs_resid_box <- ggplot(rec_diag, aes(biomass_class, abs_rec_resid_om)) +
  geom_boxplot() +
  geom_jitter(width = 0.1, size = 2, alpha = 0.7) +
  theme_bw() +
  labs(
    title = "Absolute recruitment residuals by biomass class",
    x = "Biomass class",
    y = "|Recruitment residual|"
  )

p_abs_resid_ssb
p_abs_resid_box

save_plot(p_abs_resid_ssb, "06b_absolute_OM_residuals_vs_SSB.png")
save_plot(p_abs_resid_box, "06c_absolute_residuals_by_biomass_class.png")
```

# Quarterly spawner age-structure diagnostics

This section evaluates whether the age structure of the spawning biomass in Q2 is associated with Q3 recruitment residuals. This is useful to explore  spawner-age effects in the recruitment process. 

Recruitment residuals from the Beverton–Holt stock–recruitment relationship were compared with three descriptors of the spawning stock age structure in quarter 2: mean spawner age, the proportion of spawning biomass contributed by age-1 fish, and the proportion contributed by fish aged 2 years and older.

Linear regressions showed no statistically significant association between recruitment residuals and any of the three age-structure metrics (Table X, Figure X). The estimated slope for mean spawner age was −0.46 (SE = 0.72, p = 0.53). Similarly, no relationship was detected with the proportion of spawning biomass contributed by age-1 fish (slope = 0.39, SE = 0.77, p = 0.62) or by fish aged 2 years and older (slope = −0.39, SE = 0.77, p = 0.62).

Scatterplots showed recruitment residuals centred around zero across the observed range of all three explanatory variables, with no evident linear trend or systematic change in residual variability.

```{r helper-extract-flquant}
get_flq <- function(x, slot_name) {

  obj <- slot(x, slot_name)

  if (inherits(obj, "FLQuant")) {
    return(obj)
  }

  if (inherits(obj, "FLQuants") || is.list(obj)) {
    return(obj[[1]])
  }

  if (isS4(obj) && ".Data" %in% slotNames(obj)) {
    return(obj@.Data[[1]])
  }

  stop("Could not extract FLQuant from slot: ", slot_name)
}
```



```{r q2-spawner-age-structure}
n_all   <- get_flq(ane_biol, "n")
wt_all  <- get_flq(ane_biol, "wt")
mat_all <- get_flq(ane_biol, "mat")

n_q2   <- n_all[, ac(first_yr:last_obs_yr), , ss_ssb, drop = FALSE]
wt_q2  <- wt_all[, ac(first_yr:last_obs_yr), , ss_ssb, drop = FALSE]
mat_q2 <- mat_all[, ac(first_yr:last_obs_yr), , ss_ssb, drop = FALSE]

ssb_age_q2 <- n_q2 * wt_q2 * mat_q2

ssb_age_df <- as.data.frame(ssb_age_q2) |>
  as_tibble() |>
  rename(ssb_age_q2 = data) |>
  mutate(
    age = as.numeric(as.character(age)),
    year = as.numeric(as.character(year))
  )

spawner_structure_q2 <- ssb_age_df |>
  group_by(year) |>
  mutate(
    ssb_q2_total = sum(ssb_age_q2, na.rm = TRUE),
    prop_ssb_q2 = ssb_age_q2 / ssb_q2_total
  ) |>
  summarise(
    mean_spawner_age_q2 = sum(age * prop_ssb_q2, na.rm = TRUE),
    prop_ssb_q2_age1 = sum(prop_ssb_q2[age == 1], na.rm = TRUE),
    prop_ssb_q2_age2plus = sum(prop_ssb_q2[age >= 2], na.rm = TRUE),
    ssb_q2_total = first(ssb_q2_total),
    .groups = "drop"
  ) |>
  left_join(
    rec_diag |>
      select(year, rec_obs, rec_pred_om, rec_resid_om, rec_mult_om, ssb),
    by = "year")

write_csv(spawner_structure_q2,file.path(tab_dir, "33_spawner_age_structure_Q2_by_year.csv"))

# spawner_structure_q2
```

#  Spawner age-structure diagnostics

This analysis tests whether Q3 recruitment residuals are related to the Q2 spawner age structure.

```{r maternal-effects-tests}
maternal_tests_q2 <- bind_rows(
  broom::tidy(
    lm(rec_resid_om ~ mean_spawner_age_q2, data = spawner_structure_q2)) |>
    mutate(model = "Q3 recruitment residual ~ Q2 mean spawner age"),

  broom::tidy(
    lm(rec_resid_om ~ prop_ssb_q2_age1, data = spawner_structure_q2)) |>
    mutate(model = "Q3 recruitment residual ~ Q2 proportion SSB age 1"),

  broom::tidy(
    lm(rec_resid_om ~ prop_ssb_q2_age2plus, data = spawner_structure_q2)) |>
    mutate(model = "Q3 recruitment residual ~ Q2 proportion SSB age 2+"))

write_csv(maternal_tests_q2,file.path(tab_dir, "34_maternal_effects_Q2_spawners_Q3_recruitment.csv"))

# maternal_tests_q2
```



```{r plot-mean-spawner-age-q2}
## Q3 recruitment residuals vs Q2 mean spawner age
p_mean_age_resid_q2 <- ggplot(spawner_structure_q2, aes(mean_spawner_age_q2, rec_resid_om)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(size = 2) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.8) +
  theme_bw() +
  labs(title = "Q3 recruitment residuals vs Q2 mean spawner age",
    subtitle = "Quarterly diagnostic: SSB_Q2(year t) -> Recruitment_Q3(year t)",
    x = "Mean spawner age in Q2", y = "Q3 recruitment residual")

save_plot(p_mean_age_resid_q2,"19_Q3_recruitment_residuals_vs_Q2_mean_spawner_age.png")

```

```{r plot-prop-age1-q2}
## Q3 recruitment residuals vs proportion of Q2 SSB from age 1
p_prop_age1_resid_q2 <- ggplot(spawner_structure_q2, aes(prop_ssb_q2_age1, rec_resid_om)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(size = 2) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.8) +
  theme_bw() +
  labs(title = "Q3 recruitment residuals vs Q2 proportion of SSB from age 1",
    subtitle = "Quarterly diagnostic: young-spawner contribution in Q2",
    x = "Proportion of Q2 SSB from age 1", y = "Q3 recruitment residual")

save_plot(p_prop_age1_resid_q2,"20_Q3_recruitment_residuals_vs_Q2_prop_SSB_age1.png")
```

```{r plot-prop-age2plus-q2}
## Q3 recruitment residuals vs proportion of Q2 SSB from age 2+
p_prop_age2plus_resid_q2 <- ggplot(spawner_structure_q2, aes(prop_ssb_q2_age2plus, rec_resid_om)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(size = 2) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.8) +
  theme_bw() +
  labs(title = "Q3 recruitment residuals vs Q2 proportion of SSB from age 2+",
    subtitle = "Quarterly diagnostic: older-spawner contribution in Q2",
    x = "Proportion of Q2 SSB from age 2+",y = "Q3 recruitment residual")

save_plot(p_prop_age2plus_resid_q2,"21_Q3_recruitment_residuals_vs_Q2_prop_SSB_age2plus.png")
```


```{r fig.height=9,fig.width=6}
age_plot <- p_mean_age_resid_q2/p_prop_age1_resid_q2 /p_prop_age2plus_resid_q2 +
  plot_annotation(title = "",
    theme = theme(
      plot.title = element_text(face = "bold", size = 15),
      plot.subtitle = element_text(size = 10)))
age_plot
```

# Quarterly cohort strength diagnostics

This section classifies Q3 recruitment cohorts as strong, weak or average based on standardized log recruitment.

```{r cohort-strength-q3}
cohort_strength_q3 <- rec_diag |>
  mutate(
    cohort = year,
    recruitment_season = ss_rec,
    spawning_season = ss_ssb,
    log_rec_q3 = log(rec_obs),
    log_rec_q3_anom = as.numeric(scale(log_rec_q3)),
    cohort_class = case_when(
      log_rec_q3_anom >= 1.0  ~ "Strong Q3 cohort",
      log_rec_q3_anom <= -1.0 ~ "Weak Q3 cohort",
      TRUE ~ "Average Q3 cohort"
    )
  ) |>
  select(
    cohort,
    year,
    spawning_season,
    recruitment_season,
    ssb_q2 = ssb,
    rec_q3 = rec_obs,
    rec_pred_om_q3 = rec_pred_om,
    rec_resid_om_q3 = rec_resid_om,
    rec_mult_om_q3 = rec_mult_om,
    log_rec_q3_anom,
    cohort_class
  )

cohort_summary_q3 <- cohort_strength_q3 |>
  group_by(cohort_class) |>
  summarise(
    n =  dplyr::n(),
    mean_rec_q3 = mean(rec_q3, na.rm = TRUE),
    geomean_rec_q3 = exp(mean(log(rec_q3), na.rm = TRUE)),
    mean_resid_q3 = mean(rec_resid_om_q3, na.rm = TRUE),
    median_resid_q3 = median(rec_resid_om_q3, na.rm = TRUE),
    mean_ssb_q2 = mean(ssb_q2, na.rm = TRUE),
    median_ssb_q2 = median(ssb_q2, na.rm = TRUE),
    .groups = "drop"
  )

write_csv(cohort_strength_q3,file.path(tab_dir, "35_Q3_cohort_strength_classification.csv"))

write_csv(cohort_summary_q3, file.path(tab_dir, "36_Q3_cohort_strength_summary.csv"))

cohort_summary_q3
```

## Historical Q3 cohort strength

```{r plot-cohort-strength-q3}
p_cohort_strength_q3 <- ggplot(cohort_strength_q3, aes(year, rec_q3)) +
  geom_line(alpha = 0.4) +
  geom_point(aes(shape = cohort_class), size = 2.3) +
  theme_bw() +
  labs(title = "Historical recruitment cohorts in Q3",
    subtitle = "Strong and weak cohorts based on standardized log recruitment in Q3",
    x = "Cohort year", y = "Recruitment in Q3", shape = "Cohort class")

save_plot(p_cohort_strength_q3,"21_Q3_cohort_strength_classification.png")
```



```{r plot-cohort-residuals-q3}
## Q3 recruitment residuals by cohort class
p_cohort_resid_q3 <- ggplot(cohort_strength_q3, aes(year, rec_resid_om_q3)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_line(alpha = 0.4) +
  geom_point(aes(shape = cohort_class), size = 2.3) +
  theme_bw() +
  labs(
    title = "Q3 recruitment residuals by cohort",
    subtitle = "Strong and weak Q3 cohorts relative to the fitted BH relationship",
    x = "Cohort year",
    y = "Q3 recruitment residual",
    shape = "Cohort class")

save_plot(p_cohort_resid_q3,"22_Q3_recruitment_residuals_by_cohort_strength.png")
```

# Relationship between Q3 cohort strength and Q2 SSB

This analysis evaluates whether strong or weak Q3 cohorts are associated with the spawning biomass observed in Q2 of the same year.

```{r cohort-strength-vs-ssb}
cohort_ssb_q2_summary <- cohort_strength_q3 |> group_by(cohort_class) |>
  summarise(
    n = dplyr::n(),
    mean_ssb_q2 = mean(ssb_q2, na.rm = TRUE),
    median_ssb_q2 = median(ssb_q2, na.rm = TRUE),
    min_ssb_q2 = min(ssb_q2, na.rm = TRUE),
    max_ssb_q2 = max(ssb_q2, na.rm = TRUE),
    .groups = "drop")

cohort_ssb_q2_lm <- lm(log_rec_q3_anom ~ ssb_q2, data = cohort_strength_q3)

cohort_ssb_q2_lm_table <- broom::tidy(cohort_ssb_q2_lm) |>
  mutate(
    model = "Standardized log recruitment Q3 ~ SSB Q2",
    interpretation = case_when(
      term == "ssb_q2" & p.value < 0.05 & estimate > 0 ~
        "Positive association between Q2 SSB and Q3 cohort strength",
      term == "ssb_q2" & p.value < 0.05 & estimate < 0 ~
        "Negative association between Q2 SSB and Q3 cohort strength",
      term == "ssb_q2" & p.value >= 0.05 ~
        "No clear linear association between Q2 SSB and Q3 cohort strength",
      TRUE ~ NA_character_))

write_csv(cohort_ssb_q2_summary, file.path(tab_dir, "37_Q3_cohort_strength_vs_Q2_SSB_summary.csv"))
write_csv(cohort_ssb_q2_lm_table, file.path(tab_dir, "38_Q3_cohort_strength_vs_Q2_SSB_lm.csv"))

cohort_ssb_q2_summary
cohort_ssb_q2_lm_table
```

```{r plot-cohort-strength-vs-ssb}
p_cohort_ssb_q2 <- ggplot(cohort_strength_q3, aes(ssb_q2, log_rec_q3_anom)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = Blim, linetype = "dotted") +
  geom_vline(xintercept = Bpa, linetype = "dashed") +
  geom_point(aes(shape = cohort_class), size = 2.3) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.8) +
  theme_bw() +
  labs(title = "Q3 cohort strength vs Q2 SSB",
    subtitle = "Quarterly diagnostic: SSB_Q2(year t) -> Recruitment_Q3(year t)",
    x = "SSB in Q2", y = "Standardized log recruitment in Q3", shape = "Cohort class")

save_plot(p_cohort_ssb_q2,"23_Q3_cohort_strength_vs_Q2_SSB.png")
```

```{r}
age_plot <- p_cohort_strength_q3/p_cohort_resid_q3 /p_cohort_ssb_q2 +
  plot_annotation(title = "",
    theme = theme(
      plot.title = element_text(face = "bold", size = 15),
      plot.subtitle = element_text(size = 10)))
age_plot
```


```{r recruitment-scenarios-table, echo=FALSE}

sr_scenarios <- tibble::tibble(
  Scenario = c("hist", "SS3", "s05", "s07", "extreme"),
  Description = c(
    "σR estimated from the historical residuals of the Beverton–Holt stock–recruitment model (σR ≈ 0.28).",
    "σR adopted from the SS3 stock assessment (σR = 0.33).",
    "Increased recruitment variability scenario (σR = 0.50).",
    "High recruitment variability scenario (σR = 0.70).",
    "Historical σR combined with occasional extreme low-recruitment events."
  )
)

knitr::kable(
  sr_scenarios,
  caption = "Recruitment uncertainty scenarios evaluated in the Operating Model.",
  align = c("l", "l")
)
```







```{r bootstrap-bh, eval=FALSE}
# 24. Bootstrap stability of Beverton-Holt parameters
nboot <- 1000
boot_pars <- vector("list", nboot)

for (i in seq_len(nboot)) {

  idx <- sample(seq_len(nrow(sr_data)), replace = TRUE)

  tmp <- sr_data[idx, ] |>
    arrange(year) |>
    mutate(boot_year = seq_len(n()))

  rec_boot <- FLQuant(
    tmp$rec,
    dimnames = list(
      age = "all",
      year = as.character(tmp$boot_year),
      unit = "unique",
      season = "all",
      area = "unique",
      iter = "1"
    )
  )

  ssb_boot <- FLQuant(
    tmp$ssb,
    dimnames = list(
      age = "all",
      year = as.character(tmp$boot_year),
      unit = "unique",
      season = "all",
      area = "unique",
      iter = "1"
    )
  )

  fit <- try({
    mod <- FLSR(rec = rec_boot, ssb = ssb_boot, model = bevholt)
    fmle(mod)
  }, silent = TRUE)

  if (!inherits(fit, "try-error")) {
    pars <- as.numeric(params(fit))
    names(pars) <- dimnames(params(fit))$params

    boot_pars[[i]] <- tibble(
      iter = i,
      a = pars["a"],
      b = pars["b"],
      logLik = as.numeric(logLik(fit)),
      AIC = AIC(fit)
    )
  }
}

boot_pars <- bind_rows(boot_pars) |>
  filter(is.finite(a), is.finite(b), a > 0, b > 0)

boot_summary <- boot_pars |>
  summarise(
    n_success = n(),
    success_rate = n() / nboot,
    a_median = median(a, na.rm = TRUE),
    a_q05 = quantile(a, 0.05, na.rm = TRUE),
    a_q95 = quantile(a, 0.95, na.rm = TRUE),
    b_median = median(b, na.rm = TRUE),
    b_q05 = quantile(b, 0.05, na.rm = TRUE),
    b_q95 = quantile(b, 0.95, na.rm = TRUE),
    b_q25 = quantile(b, 0.25, na.rm = TRUE),
    b_q75 = quantile(b, 0.75, na.rm = TRUE)
  )

write_csv(boot_pars, file.path(tab_dir, "22_bootstrap_BH_parameters.csv"))
write_csv(boot_summary, file.path(tab_dir, "23_bootstrap_BH_summary.csv"))

boot_summary
```

```{r plot-bootstrap, eval=FALSE}
p_boot_b <- ggplot(boot_pars, aes(x = b)) +
  geom_histogram(bins = 40) +
  geom_vline(xintercept = b_om, linetype = "dashed") +
  theme_bw() +
  labs(
    title = "Bootstrap distribution of Beverton-Holt b parameter",
    subtitle = paste0("Successful fits: ", nrow(boot_pars), " / ", nboot),
    x = "BH b parameter",
    y = "Frequency"
  )

p_boot_ab <- ggplot(boot_pars, aes(a, b)) +
  geom_point(alpha = 0.35, size = 1.5) +
  geom_point(aes(x = a_om, y = b_om), size = 3) +
  theme_bw() +
  labs(
    title = "Bootstrap correlation between Beverton-Holt parameters",
    subtitle = "Diagnostic for parameter uncertainty and identifiability",
    x = "BH a parameter",
    y = "BH b parameter"
  )

p_boot_b
p_boot_ab

save_plot(p_boot_b, "14_bootstrap_BH_b_parameter.png")
save_plot(p_boot_ab, "17_bootstrap_BH_a_b_correlation.png")
```



```{r decision-table, eval=FALSE}
# 25. Final conditioning decision table
conditioning_decisions_quarterly <- tibble(
  question = c(
    "Should a stock-recruitment relationship be used?",
    "Which SR relationship should be used in the base OM?",
    "Should Ricker overcompensation be used?",
    "Should recruitment deviations be lognormal?",
    "Should recruitment deviations be autocorrelated?",
    "Should productivity regimes be included?",
    "Should biomass-dependent recruitment variance be included?",
    "Should environmental covariates be included?",
    "Should maternal or spawner-age effects be included?",
    "Should sporadic recruitment events be included?",
    "Should SR parameter uncertainty be explored?"
  ),
  quarterly_diagnostic = c(
    "Q2 SSB linked to Q3 recruitment",
    "BH vs segmented vs Ricker comparison using Q2 SSB and Q3 recruitment",
    "Ricker comparison and recruitment at high Q2 SSB",
    "Q3 recruitment residual distribution, QQ plot, Shapiro-Wilk and tail diagnostics",
    "ACF, PACF, Ljung-Box, AR(1), runs test on Q3 recruitment residuals",
    "Changepoint, breakpoints and productivity diagnostics using Q3 recruitment",
    "Absolute Q3 recruitment residuals vs Q2 SSB, biomass-class tests",
    "No tested mechanistic environmental driver",
    "Q3 recruitment residuals vs Q2 spawner age structure",
    "Extreme Q3 residual years and Q3 cohort strength classification",
    "Bootstrap BH parameters using Q2 SSB and Q3 recruitment"
  ),
  base_OM_decision = c(
    "Yes",
    "Beverton-Holt",
    "No",
    "Yes",
    "No",
    "No",
    "No",
    "No",
    "No, unless diagnostics show a clear relationship",
    "No as explicit process; retained through stochastic recruitment variability",
    "Not in base OM; use sensitivity scenarios"
  )
)

write_csv(conditioning_decisions_quarterly,file.path(tab_dir,"39_conditioning_decision_table_quarterly.csv"))

conditioning_decisions_quarterly
```



# Definir cuál era el rango biológicamente razonable de Fcap

